%%%%%%%%%%%%%%%%%%%%%%%%%%%% Document Setup %%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[10pt]{article}

% This is a helpful package that puts math inside length specifications
\usepackage{calc}

% Layout: Puts the section titles on left side of page
\reversemarginpar

%% Set up for letter-sized paper
\usepackage[paper=letterpaper,
marginparwidth=1.1In,     % Length of section titles
marginparsep=0in,       % Space between titles and text
margin=1in,               % 1 inch margins
includemp]{geometry}

%% Get rid of indenting throughout entire document
\setlength{\parindent}{0in}

%% Make smaller skip between paragraphs
\setlength{\parskip}{-0.1in}

%% Reference the last page in the page number
\usepackage{fancyhdr,lastpage}
\pagestyle{fancy}
\fancyhf{}\renewcommand{\headrulewidth}{0pt}
\fancyfootoffset{\marginparsep+\marginparwidth}
\newlength{\footpageshift}
\setlength{\footpageshift}
{0.5\textwidth+0.5\marginparsep+0.5\marginparwidth-2in}
\lfoot{\hspace{\footpageshift}%
  \parbox{4in}{\, \hfill %
    \arabic{page} of \protect\pageref*{LastPage}
    \hfill \,}}

%% PDF bookmarks
\usepackage{color,hyperref}
\definecolor{darkblue}{rgb}{0.0,0.0,0.3}
\hypersetup{colorlinks,breaklinks,
            linkcolor=darkblue,urlcolor=darkblue,
            anchorcolor=darkblue,citecolor=darkblue}
% for URLs
\usepackage{verbatim}

%%%%%%%%%%%%%%%%%%%%%%%% End Document Setup %%%%%%%%%%%%%%%%%%%%%%%%%%%%


%%%%%%%%%%%%%%%%%%%%%%%%%%% Helper Commands %%%%%%%%%%%%%%%%%%%%%%%%%%%%

% The title (name) with a horizontal rule under it
%
% Usage: \makeheading{name}
%
% Place at top of document. It should be the first thing.

\newcommand{\makeheading}[1]{
  \hspace*{-\marginparsep minus \marginparwidth}%
  \begin{minipage}[t]{\textwidth+\marginparwidth+\marginparsep}%
    \begin{center}
      {\large \bfseries #1}\\[-0.15\baselineskip]%      
      \rule{\columnwidth}{1pt}%
    \end{center}
  \end{minipage}
}

%% The section headings
%% Follow this section IMMEDIATELY with the first line of the section
%% text. Do not put whitespace in between.
\renewcommand{\section}[2]%
        {\pagebreak[3]\vspace{1.3\baselineskip}%
         \phantomsection\addcontentsline{toc}{section}{#1}%
         \hspace{0in}%
         \marginpar{
         \raggedright \scshape #1}#2}

%% tabular styles for including dates in list of CV content
\usepackage{array, xcolor}
\definecolor{lightgray}{gray}{0.8}
\newcolumntype{L}{p{0.19\textwidth}}
\newcolumntype{R}{p{0.81\textwidth}}
\newcommand\VRule{\color{lightgray}\vrule width 0.5pt}

%% To compress itemized environments
\usepackage{enumitem}
\setlist{nolistsep}
\setlist{noitemsep}


%% To add some paragraph space between lines.
%% This also tells LaTeX to preferably break a page on one of these gaps
%% if there is a needed pagebreak nearby.
\newcommand{\blankline}{\quad\pagebreak[3]}
\newcommand{\halfblankline}{\quad\vspace{-0.5\baselineskip}\pagebreak[3]}


%% NOTE: \rcollength is the width of the right column of the table
\newlength{\rcollength}\setlength{\rcollength}{2.75in}%


%% THIS IS THE PRIMARY COMMAND USED FOR ENTERING CV DATA!!!!
\newcommand{\cventry}[2]{%
\begin{tabular}[t]{R !{\VRule} L}
  \begin{minipage}[t][][t]{0.81\textwidth}
    #2
  \end{minipage}
  &
  \parbox[t][][t]{0.19\textwidth}{#1}
\end{tabular}%
\vspace{0.25em}
}


%%%%%%%%%%%%%%%%%%%%%%%% End Helper Commands %%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%% Begin CV Document %%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
% \makeauthorbold{Rominger}
\makeheading{Andrew J.~Rominger}

\section{Contact Information}
%
\begin{tabular}[t]{p{\textwidth-\rcollength}p{\rcollength}}
  \multicolumn{2}{l}{Department of Environmental Science, Policy \&
    Management} \\  
  University of California, Berkeley &\textit{Phone:} +1-650-862-6063 \\
  130 Mulford Hall                          &\textit{E-mail:} \verb|rominger@berkeley.edu|\\
  Berkeley, CA 94720 USA               & \textit{Web:} \verb|nature.berkeley.edu/~rominger|\\
\end{tabular}

\section{(A) Professional Preparation}
%
\cventry{2016}{
  {\bf University of California, Berkeley} \\
  Ph.D. Candidate in Environmental Science, Policy \& Management \\
  {\it Committee:} Rosemary Gillespie, John Harte \& Charles Marshall
}
%
\cventry{2009}{
  {\bf Stanford University} \\
   B.S. in Biological Sciences \\
  {\it Advisors:} Elizabeth Hadly \& Rodolfo Dirzo \\
  {\it Honors thesis:} Both neutral and deterministic processes drive
  community structure
}

\section{(B) Appointments}
%
\cventry{2016}{
  {\bf Post Doctoral Scholar,} UC Berkeley \\
  Berkeley Initiative in Global Change Biology
}
%
\cventry{2014}{
  {\bf Graduate Student Instructor,} UC Berkeley \\
  ESPM 174: Design and
  Analysis of Ecological Studies \\
  {\it Instructor:} Perry de Valpine \\
  {\bf Graduate Student Instructor,} UC Berkeley \\
  INTEGRATIVE BIOLOGY 166: Evolutionary Biogeography \\
  {\it Instructor:} Anthony Barnosky
}
%
\cventry{2011--present}{
  {\bf NSF Graduate Research Fellow,} UC Berkeley \\
  Building ecological theories of abundance, energetics and spatial
  distribution that incorporate phylogenetic information. Studying the
  evolution, biogeography and ecology of Hawaiian arthropods. Building
  statistical methods that can leverage large amounts of newly
  digitized biodiversity data and gain insights into how ecological
  communities change in response global change.  }
%
\cventry{2010}{
  {\bf Fulbright Scholar,} Pontificia Universidad Cat\'olica de Chile \\
  Used databases of fossil collections to study the evolution Phanerozoic
  biodiversity. Developed models of diversification using principles from
  thermodynamics, evolutionary niche construction and queuing processes. \\
  {\it Host:} Pablo Marquet
}
%
\cventry{2009}{
  {\bf Teaching Assistant,} Stanford University \\
  BIOLOGY 121: Biogeography \\
  {\it Instructor:} Elizabeth Hadly
}
%
\cventry{2008}{
  {\bf Data Analyst,} Stanford University \\
  Analyzed herbaria records to detect signs of geographic
  range shift under $20^{th}$ Century climate change. \\
  \\[-0.75em]
  {\bf Intern in Paleobiology,} Smithsonian Institution \\
  Used Smithsonian collection to reconstruct phylogenetic
  relationships of Permian brachiopods and
  studied evolution of theoretical morphospace occupancy. \\
  {\it Mentors:} Doug Erwin, Peter Wagner }


\section{(C) Products (i)}
%
\cventry{in press}{
  \begin{itemize}
  \item[]\hspace{-1.1\leftmargin} {\bf Rominger AJ}, Merow
    C. (2016). {\tt meteR}: An {\tt R} package for testing the Maximum
    Entropy Theory of Ecology. In press at {\it Methods in Ecology and
      Evolution}.
  \end{itemize}
}
%
\cventry{2016}{
  \begin{itemize}
  \item[]\hspace{-1.1\leftmargin} {\bf Rominger AJ}, {\it et
      al.}. (2016). Community assembly on isolated islands: Macroecology meets
    evolution. {\it Global Ecology and Biogeography} 25: 769--780.
  \end{itemize}
}
%
\cventry{2015}{
  \begin{itemize}
  \item[]\hspace{-1.1\leftmargin} Harte J, {\bf Rominger AJ}, Zhang
    W. (2015). Integrating macroecological metrics and community taxonomic
    structure. {\it Ecology Letters} 18: 1068--1077.
  \end{itemize}
}
%
\cventry{2012}{
  \begin{itemize}
  \item[]\hspace{-1.1\leftmargin} Karp DS, {\bf Rominger AJ}, Zook J,
    Ranganathan J, Ehrlich PR \& Daily GC. (2012). Intensive
    agriculture erodes $\beta$-diversity
    at large scales. {\it Ecology Letters} 15: 963--970.
  \end{itemize}
}
%
\cventry{2009}{
  \begin{itemize}
  \item[]\hspace{-1.1\leftmargin} {\bf Rominger AJ}, Miller TEX \& Collins
    SL. (2009). Relative contributions of neutral and niche-based
    processes to the structure of a desert grassland grasshopper
    community. {\it Oecologia} 161: 791--800.
  \end{itemize}
}

\section{(C) Products (ii)}
%
\cventry{in press}{
  \begin{itemize}
  \item[]\hspace{-1.1\leftmargin} Sardinas HS, Tom K, {\bf Rominger
      AJ}, Kremen C. (2015). Patterns of native bee crop pollination within
    agricultural fields are limited by nest site location. In press at
    {\it Ecological Applications}.
  \end{itemize}
}
%
\cventry{2013}{
  \begin{itemize}
  \item[]\hspace{-1.1\leftmargin} Harte J, Kitzes J, Newman E, {\bf
      Rominger AJ}. (2013). Taxon categories and the universal
    species-area relationship: A comment on Sizling et al.  {\it The
      American Naturalist} 181: 282--287.
  \end{itemize}
}
%
\cventry{2012}{
  \begin{itemize}
  \item[]\hspace{-1.1\leftmargin} Maurer BA, Kembel SW, {\bf Rominger
      AJ} \& McGill BJ. (2012). Estimating metacommunity extent using
    data on species abundances, environmental variation, and
    phylogenetic relationships across geographic space. {\it
      Ecological Informatics} 13: 114--122.
  \end{itemize}
}

\section{(D) Synergistic Activities}
%
\cventry{2013--present}{
  {\bf Board Member} \\
  Talking Talons Youth Leadership Community Fund, an organization that
  funds environmental education projects.
}
%
\cventry{2009--present}{
  {\bf Community presentation speaker} \\
  Present at youth and environmental group meetings including Central
  New Mexico Audubon Society and Pacific Internship Programs for
  Exploring Science about science, conservation and environmental
  education.
}
%
\cventry{2009}{
  {\bf Splash instructor} \\
  Thought an interactive course about global change biology to K-12
  students as part of Stanford University's Educational Studies Splash Program.
}
%
\cventry{2006--present}{
  {\bf Natural history docent} \\
  Lead classroom and community tours of Jasper Ridge Biological
  Preserve (Stanford University) focusing on local
  conservation issues, ecology, evolution and geology.
}

% \section{Collaborators}
% %
% \begin{tabular}[t]{p{1\textwidth}}
%   E.~Armstrong (U Hawaii); 
%   L.~Becking (Institute for Marine Resources and Ecosystem Studies); 
%   G.~Bennett (U of Hawaii); 
%   M.~Brewer (UC Berkeley);
%   S.~Collins (U New Mexico); 
%   D.~Cotoras (UC Santa Cruz);
%   G.~Daily (Stanford University);
%   R.~Dirzo (Stanford University);
%   P.~Ehrlich (Stanford University);
%   D.~Erwin (Smithsonian);
%   C.~Ewing (U Hawaii);
%   M.~Fuentes (Consejo Nacional de Investigaciones Cient\'ificas y
%   T\'ecnicas);
%   R.~Gillespie (UC Berkeley);
%   K.~Goodman (UC Berkeley);
%   D.~Gruner (U Maryland);
%   E.~Hadly (Stanford University);
%   J.~Harte (UC Berkeley);
%   S.~Kembel (U Qu\'ebec \`a Montr\'eal);
%   J.~Lim (UC Berkeley);
%   P.~Marquet (Pontificia Universidad Cat\'olica de Chile);
%   C.~Marshall (UC Berkeley);
%   N.~Martinez (U Arizona);
%   B.~Mauer (Michigan State University);
%   B.~McGill (U Maine);
%   C.~Merow (U Connecticut);
%   T.~Miller (Rice University);
%   J.~O'Dwyer (U Illinois Urbana-Champaign);
%   P.~O'Grady (UC Berkeley);
%   D.~Percy (Natural History Museum, London);
%   D.~Price (U Hawaii);
%   G.~Roderick (UC Berkeley);
%   K.~Shaw (Cornell University);
%   F.~Valdovinos (U Arizona);
%   P.~Wagner (Smithsonian).
%   X.~Xiao (U Main)
% \end{tabular}



\end{document}
